---
output: 
  github_document:
    html_preview: true
    fig_height: 6
    fig_width: 6
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

**IMPORTANT MESSAGE:** this package is new and hasn't yet been extensively tested. Use due diligence when handling sensitive or confidential data.

# Installation

'deidentifyr' isn't on CRAN yet. You can install it from github with 'devtools':

```{r, eval = FALSE}
devtools::install_github('wilkox/deidentifyr')
```

# Walkthrough

Here's an example dataset containing some patient data.

```{r}
set.seed(1)
n <- 10
MRNs <- sample(10000000:99999999, n)
DOBs <- lubridate::today() - lubridate::dyears(sample(18:99, n, replace = T))
times_in_hospital <- sample(1:100, n, replace = T)
patient_data <- data.frame(MRN = MRNs, DOB = DOBs, 
                           time_in_hospital = times_in_hospital)
patient_data
```

There are two variables in this data frame containing personally identifying information: `MRN` and `DOB`. We could just remove these columns and generate a random ID number for each patient, but that would make it difficult to match the patients if we wanted to merge two data frames together. The solution is to generate a unique ID code, a cryptographic hash (SHA-256), from the identifying columns. This type of hash has two useful properties: it is very unlikely that the same hash would be generated for two people who have different information; and it is near impossible to recover the personal information from the hash. For most datasets, using only the first ten characters of the hash is sufficient to generate unique IDs.

We can generate these unique IDs with the `deidentify()` function. The first argument to `deidentify()` is the data frame, and after that we can list the columns from which to generate the IDs.

```{r}
library(deidentifyr)
patient_data <- deidentify(patient_data, MRN, DOB)
patient_data
```

The `MRN` and `DOB` columns have been removed, and replaced with a new column called `id` containing a unique hash for each patient. If you don't want to remove the original columns, `deidentify()` can be called with the argument `drop = FALSE`. You can also choose a different name for the ID column with `key = "name"`.

The same identifying details will always generate the same hash. This means that a different data frame deidentified in the same way will have the same IDs for each patient.

```{r}
sexes <- sample(c("F", "M"), n, replace = T)
patient_data2 <- data.frame(MRN = MRNs, DOB = DOBs, 
                           sex = sexes)
patient_data2
patient_data2 <- deidentify(patient_data2, DOB, MRN)
patient_data2
```

Note that it didn't matter that we listed the identifying columns in a different order the second time we called `deidentify()`. We can now match patients between the data frames without needing to reidentify them.

```{r}
combined_data <- merge(patient_data, patient_data2, by = "id")
combined_data
```

# Similar packages

* [anonymizer](https://github.com/paulhendricks/anonymizer) takes a slightly different approach to the same problem.
