---
output: 
  github_document:
    html_preview: true
    fig_height: 6
    fig_width: 6
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# Important message

This package is still under development and hasn't yet been extensively tested. Use due diligence when handling sensitive or confidential data.

# Installation

'deidentifyr' isn't on CRAN yet. You can install it from github with 'devtools':

```{r, eval = FALSE}
devtools::install_github('wilkox/deidentifyr')
```

# Walkthrough

Here's an example dataset containing some patient data.

```{r, include = FALSE}
set.seed(1)
n <- 10
MRNs <- sample(10000000:99999999, n)
DOBs <- lubridate::today() - lubridate::dyears(sample(18:99, n, replace = T))
days_in_hospitals <- sample(1:100, n, replace = T)
patient_data <- data.frame(MRN = MRNs, DOB = DOBs, 
                           days_in_hospital = days_in_hospitals)

```{r}
patient_data
```

There are two variables in this data frame containing personally identifying information: `MRN` and `DOB`. We could just remove these columns and generate a random ID number for each patient, but that would make it difficult to match the patients if we wanted to merge two data frames together. The solution is to generate a unique ID code, a cryptographic hash (SHA-256), from the identifying columns. This type of hash has two useful properties: it is very unlikely that the same hash would be generated for two people who have different information; and it is near impossible to recover the personal information from the hash (though in certain circumstances it might be very easy; see [Salting](#salting), below). For most datasets, using only the first ten characters of the hash is sufficient to generate unique IDs.

We can generate these unique IDs with the `deidentify()` function. The first argument to `deidentify()` is the data frame, and after that we can list the columns from which to generate the IDs.

```{r}
library(deidentifyr)
patient_data <- deidentify(patient_data, MRN, DOB)
patient_data
```

The `MRN` and `DOB` columns have been removed, and replaced with a new column called `id` containing a unique hash for each patient. If you don't want to remove the original columns, `deidentify()` can be called with the argument `drop = FALSE`. You can also choose a different name for the ID column with `key = "name"`.

The same identifying details will always generate the same hash. This means that a different data frame deidentified in the same way will have the same IDs for each patient.

```{r, include = FALSE}
sexes <- sample(c("F", "M"), n, replace = T)
patient_data2 <- data.frame(MRN = MRNs, DOB = DOBs, 
                           sex = sexes)
```

```{r}
patient_data2
patient_data2 <- deidentify(patient_data2, DOB, MRN)
patient_data2
```

Note that it didn't matter that we listed the identifying columns in a different order the second time we called `deidentify()`. We can now match patients between the data frames without needing to reidentify them.

```{r}
combined_data <- merge(patient_data, patient_data2, by = "id")
combined_data
```

## Salting

In certain circumstances, there is a potential method somebody could use to reidentify the patients. Suppose a bad actor happened to have access to a master list of all the patients who had ever been admitted to the hospital. If they could guess which columns (i.e. which pieces of personally identifying information) you used to create the unique IDs, they could generate IDs from this master list using the same hashing method. They could then compare their IDs to the ones you created, and figure out who each patient is in the deidentified dataset.

A solution to this is to add an extra piece of information to the hash, which is not personally identifying information about the patients but a secret known only to you. By adding this extra piece, called a "salt", completely different unique IDs will be generated. You will be able to regenerate the same IDs from a different dataset if you wanted, by adding the same salt, so your ability to match patients between deidentified datasets will not be lost. However, unless the bad actor manages to discover your secret salt, they will not be able to generate a list of potential IDs from which to reidentify the patients. You can add a salt by calling `deidentify()` with the extra argument `salt = "mysalt"`.

Decisions about using or not using a salt, how to keep the salt a secret, and what personally identifying information to include in the hash will depend on the nature of the dataset and the context in which it will be used. Having access to confidential data, particularly about people's health, is a privilege and a responsibility. Take some time to make these decisions carefully.

# Similar packages

* [anonymizer](https://github.com/paulhendricks/anonymizer) takes a different approach to the same problem.
